---
title: "Course Project 1"
author: "Juan"
date: "22 de octubre de 2020"
output: html_document
---
##Instructions

   1. Code for reading in the dataset and/or processing the data
   2. Histogram of the total number of steps taken each day
   3. Mean and median number of steps taken each day
   4. Time series plot of the average number of steps taken
   5. The 5-minute interval that, on average, contains the maximum number of steps
   6. Code to describe and show a strategy for imputing missing data
   7. Histogram of the total number of steps taken each day after missing values are          imputed
   8. Panel plot comparing the average number of steps taken per 5-minute interval            across weekdays and weekends
   9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the         report
   
   
## Step1-  Code for reading in the dataset and/or processing the data

```{r, echo=TRUE}
setwd("C:/Users/asus/Documents/R-studio/Scripts/proyectos R/Reproducible Research Project 1")
library(lubridate)
unzip("../Reproducible Research Project 1/repdata_data_activity.zip")

activity <- read.csv("activity.csv")
activity$date <- as.POSIXct(strptime(activity$date, "%Y-%m-%d", Sys.timezone()))
weekday <- weekdays(activity$date, abbreviate = FALSE)
activity <- cbind(activity, weekday)
 
dim(activity)
names(activity)
summary(activity)
head(activity)
str(activity)

library(lubridate)
activity$date <- ymd(activity$date)
length(unique(activity$date))

```

##Step2- Histogram of the total number of steps taken each day

```{r, echo=TRUE}

library(ggplot2)

#total step per day
stepxday <- aggregate(steps ~ date, activity, sum, na.rm = TRUE)
png("plot2.png")
hist(stepxday$steps, main = "Histogram Total of Step per Day", xlab = "Total Step", col = "brown")

dev.off()

```

##Step3 Mean and median number of steps taken each day
```{r, echo=TRUE}

mean(stepxday$steps)
median(stepxday$steps)
```

##Step4 Time series plot of the average number of steps taken
```{r, echo=TRUE}

stepavrg <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
png("plot4.png")
plot(steps ~ interval, data = stepavrg, type = "l", col = "brown")

dev.off()

```

##Step5 The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo=TRUE}

Maxintervalsteps <- stepavrg[which.max(stepavrg$steps), ]$interval

Maxintervalsteps

```

##Step6 Code to describe and show a strategy for imputing missing data

using mean/median strategy
```{r, echo=TRUE}

sum(is.na(activity$steps))

meanstepxint <- function(interval) {
        stepavrg[stepavrg$interval == interval, ]$steps
}

activity_imputed <- activity
for (i in 1:nrow(activity_imputed)) {
        if(is.na(activity_imputed[i,]$steps)){
                activity_imputed[i,]$steps <- meanstepxint(activity_imputed[i,]$interval)
        }
        
}

```

##Step7  Histogram of the total number of steps taken each day after missing values are          imputed

```{r, echo=TRUE}

totalstepmv <- aggregate(steps ~ date, data = activity_imputed, sum)
png("plot7.png")
hist(totalstepmv$steps, main = "Histogram Total Steps taken each day", xlab = "Total Step for Day", col = "brown")
mean(totalstepmv$steps)
median(totalstepmv$steps)
dev.off()

```
##Step8 Panel plot comparing the average number of steps taken per 5-minute interval            across weekdays and weekends

```{r, echo=TRUE}
library(ggplot2)
library(lattice)



activity_imputed$date <- as.Date(strptime(activity_imputed$date, format = "%Y-%m-%d"))
activity_imputed$day <- weekdays(activity_imputed$date)
for (i in 1:nrow(activity_imputed)) {
        if (activity_imputed[i, ]$day %in% c("sabado", "domingo")) {
                activity_imputed[i,]$day <- "weekend"
        }
        else{
                activity_imputed[i,]$day <- "weekday"
        }
}

SbyDay <- aggregate(activity_imputed$steps ~ activity_imputed$interval + activity_imputed$day, activity_imputed, mean)

names(SbyDay) <- c("interval", "day", "steps")

png("plot8.png")

xyplot(steps ~ interval | day, SbyDay, type = "l", layout = c(1,2), xlab = "Interval",
       ylab = "Average Steps")

dev.off()

```        